---
title: "Read and Format Data"
output: html_notebook
---


Yahooo!!! 


This is just a small example of how a reproducible project might look.
Each notebook does very little (less than in a typical analysis).

This is the first notebook to evaluate.  (number 001).  Outputs from this
notebook will be used later, so we create a directory, `outputs/001` to
write outputs to.
```{r}
dir.create("outputs/001", showWarnings = FALSE, recursive = TRUE)
```

Eric has stored the standard "iris" data from R into a CSV file
in the `data` directory in this project.  We will read it in.
```{r}
library(tidyverse)

dat <- read_csv("data/iris.csv")
```

Now, we will make the column names more to our liking and change
the order to be more pleasing, and also give a unique ID to
each plant.
```{r}
dat_final <- dat %>%
  rename(
    species = Species,
    sepal_length = Sepal.Length,
    sepal_width = Sepal.Width,
    petal_length = Petal.Length,
    petal_width = Petal.Width
  ) %>%
  relocate(species, .before = sepal_length) %>%
  mutate(
    indiv = sprintf("plant%03d", 1:n()), 
    .before = species
  )
```

These changes are somewhat gratuitous.  Usually you would have much
more data cleaning/formatting to do.  This is just for an example.

With that done, we will write the object out as as RDS file for
use in later scripts or notebooks.

```{r}
write_rds(dat_final, "outputs/001/dat_final.rds")
```

## SessionInfo

At the end, you should write out the SessionInfo, which tells you
about the version of R that was used as well as information about
which packages were employed.

R comes with a command called `sessionInfo()`, which is standard,
but if you install the 'sessioninfo' package you can get output that
is more easily read.  The following code uses the `sessioninfo` package
if it is installed, and the standard `sessionInfo()` otherwise.
```{r}
if("sessioninfo" %in% rownames(installed.packages())) {
  sessioninfo::session_info()
} else {
  sessionInfo()
}
```